---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Load data into correct format

* Check if activity.csv already exists, unpack if necessary

* using read_csv import data with correct data types

* look at the head of the data and summary to figure out basics

```{r}
if (!file.exists("./activity.csv"))
{
        unzip("./activity.zip")
}
library(readr)
activity <- read_csv("activity.csv", col_types = cols(date = col_date(format = "%Y-%m-%d"), 
    steps = col_number()))
head(activity)
summary(activity)

```
Provide a guick look at steps destibution:

* by interval

* day

* both

```{r}
library(ggplot2)
qplot(activity$interval,activity$steps)
qplot(activity$date,activity$steps)
x <- ggplot(data = activity, mapping = aes(x=date, y=interval, col=steps))
x+geom_point()
```

## What is mean total number of steps taken per day?

```{r}
activityclear <- na.omit(activity)
activityclear$datenum <- as.numeric(activityclear$date)
activityclear <- activityclear[,-2]
 library(party)
 cfit1 <- ctree(steps ~ ., data = activityclear[,-3])
 activityclear$pred <- predict(cfit1,activityclear)
qplot(activityclear$interval,activityclear$pred, geom = "line") 
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
